\documentclass[12pt]{article}
\usepackage{palatino}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{latexsym} 

\usepackage{color}
\usepackage{amsmath,amssymb}
\usepackage[utf8]{inputenc}
\pagestyle{empty}

\setlength{\parindent}{0in}
\setlength{\parskip}{.1in}

\newcounter{showsolution}
\setcounter{showsolution}{0}
\newcommand{\solution}[2]{\ifnum\value{showsolution}>0{#1}\else{#2}\fi}

\newcounter{QuestionCounter}
%\newcommand{\Question}[1]{\refstepcounter{QuestionCounter} {\bf Question \arabic{QuestionCounter}}. [#1 points] }

\newcommand{\Question}[2]{\refstepcounter{QuestionCounter} \textbf{\Large {Q\arabic{QuestionCounter}}. #1\hfill\textit{#2 points}}}

\begin{document}

{\ }
\vskip 1in

\centerline{\Huge\bf First Midterm Exam}
\vskip .5in
\centerline{\LARGE\bf 601.467/667 Introduction to Human Language Technology}
\vskip .2in
\centerline{\Large Fall 2020}
\vskip .1in
\centerline{\Large Johns Hopkins University}
\vskip .1in
\centerline{\Large Co-ordinator: Philipp Koehn}
\vskip .3in
\centerline{\Large 8 October 2020}
\vskip 1in
\centerline{\Large Complete all questions.}
\vskip .2in
\centerline{\Large Use additional paper if needed.}
\vskip .2in
\centerline{\Large Time: 75 minutes.}
\vskip 1in
\centerline{\Large Name of student: {\rule{3in}{0.4pt}}}

\newpage
\Question{\Large Language Models}{10}

If you train a trigram language model on the collected works of Shakespeare,
explain in detail how to use the language model to generate large quantities of
new Shakespearian sounding text.


\newpage
\Question{\Large Morphology}{10}
\begin{enumerate}
\item Which type of language is likely to have longer words:

   \hspace*{3mm} {\sc fusional language} or {\sc agglutenative language} (circle one)
   \solution{}
   
\vspace*{1cm}

\item Derivational and inflectional morphology in English.

\begin{enumerate}
\item Give a highly productive derivational {\it prefix} in English: \_\_\_\_\_\_\_
\solution{}

\vspace*{0.5cm}

\item Give a highly productive derivational {\it infix} in English: \_\_\_\_\_\_\_
\solution{}

\vspace*{0.5cm}

\item Give a highly productive derivational {\it suffix} in English: \_\_\_\_\_\_\_
\solution{}

\vspace*{0.5cm}
\item Give an essentially non-productive {\it inflectional  plural suffix} in English: \_\_\_\_\_\_\_
\solution{}
    
\vspace*{0.5cm}

\end{enumerate}

\item Give an example of a word in English with more than one derivational parse,
  and illustrate {\it two} distinct derivational parses for that word via morpheme bracketing using 
  [ and ] and write the approximate meaning of each.
  Do not use an example containing the word 'lock' or 'employ'.
  \solution{}


\end{enumerate}

\newpage
\Question{\Large Syntax}{20}

\item Recall the definition of a context free grammar, $\mathcal{G}$: a tuple $(N, \Sigma, S, R)$ where

\begin{itemize}
    \item $N$ is a set of \textbf{nonterminal} symbols;
    \item $\Sigma$ is a set of \textbf{terminal} symbols, such that $N \cap \Sigma = \emptyset$;
    \item $S \in N$ is a special \textbf{start symbol}; and
    \item $R$ is a \textbf{set of rules} of the form $A\rightarrow B~C$ (for $B, C \in N$) or $A\rightarrow a$ (for $a \in \Sigma$)
\end{itemize}

\begin{enumerate}
\item (10 points) Consider the following partial grammar:
\begin{itemize}
    \item $N = \{\mathrm{S,NP,VP,PP,IN,NN,VB,DT}\}$
    \item $\Sigma = \{\mathrm{time, flies, like, an, arrow}\}$
    \item $S=$ S
\end{itemize}

Design a set of rules $R$ that can produce exactly two sensible parses of the sentence \emph{time flies like an arrow}.
The parses should correspond to the following two meanings of this sentence:

\begin{enumerate}
    \item There is a kind of insect known as a \emph{time fly} that appreciates archery
    \item One's children grow up very fast.
\end{enumerate}

Your grammar should have exactly 13 rules.

\newpage
\item (10 points) A derivation $d$ of a grammar $\mathcal{G}$ is a set of rules that recursively expand the root symbol until there are no remaining nonterminals to expand.
For example, for the grammar
%
\begin{align}
\mathcal{G} & = (N, \Sigma, S, R) \\
  & = (\{\mathrm{S,X}\}, \{\mathrm{y}\}, \mathrm{S}, \{S\rightarrow X~X, X\rightarrow X~X, X\rightarrow y\})
\end{align}
%
one possible derivation is
%
\begin{table}[h]
    \centering
\begin{tabular}{rl}
    \textbf{string} & \textbf{rule to apply} \\
    S   &  S $\rightarrow$ X~X \\
    X~X &  X $\rightarrow$ X~X \\
    X~X~X &  X $\rightarrow$ y \\
    y~X~X &  X $\rightarrow$ y \\
    y~y~X &  X $\rightarrow$ y \\
    y~y~y &  $\blacksquare$ \\
\end{tabular}
\end{table}

A \emph{weighted grammar} is one in which each rule $r \in R$ is associated with a score. The score of a derivation is then the sum of the scores of the rules in that derivation.
Assign weights to your rules in Question 1 so that the derivation above interpretation (b) above has a higher score.

\end{enumerate}

%\newpage
%\Question{\Large Semantics}{20}

\newpage
\Question{\Large Deep Learning: Sigmoid Function}{10}

\vskip .25in
The sigmoid function is defined as follows:
\begin{equation*}
    \sigma \triangleq \frac{1}{1 + e ^{-x}}
\end{equation*}

\begin{enumerate}
    \item What is $\sigma (0)$, $\lim _{x \rightarrow \infty} \sigma(x)$, and $\lim _{x \rightarrow - \infty} \sigma(x)$?
    \vskip 2.0in
%    \item Derive $\sigma'(x) = \frac{e^{-x}}{(1 + e ^{-x})^2} = \sigma(x)(1 - \sigma(x))$
%    \vskip 1.0in
%    \item Show that $\sigma'(x) \geq 0$
%    \vskip 1.0in
%    \item What is $\sigma'(0)$, $\lim _{x \rightarrow \infty} \sigma'(x)$, and $\lim _{x \rightarrow - \infty} \sigma'(x)$?
%    \vskip 0.5in
%    \item What is the range of $\sigma(x)$?
%    \vskip 0.5in
    \item Plot the sigmoid function by hand by considering the above answers.
\end{enumerate}

\newpage
\Question{Deep Learning: Softmax Function}{15}

\vskip .25in

Suppose we have a $J$ dimensional vector $\mathbf{h} \in \mathbb{R} ^J$, the softmax function for element $j$ is defined as follows:
\begin{equation*}
    [ \text{softmax} (\mathbf{h}) ]_j \triangleq \frac{e ^{h_j}}{\sum _{i=1} ^J e ^{h_i}}
\end{equation*}

\begin{enumerate}
    \item Why the softmax function is used as a probabilistic distribution function?
    Please discuss it by considering $\sum _{j=1} ^J [ \text{softmax} (\mathbf{h}) ]_j$ and the range of $[ \text{softmax} (\mathbf{h}) ]_j$.
    \vskip 2in
    \item Prove that the softmax function becomes the sigmoid function when $J=2$ by setting $h_2 - h_1 = -x$ (or $h_1 - h_2 = -x$)
\end{enumerate}

\newpage
\Question{\Large Information Retrieval}{10}


\begin{enumerate}
\item Draw an inverted index data structure for the following eight documents, so that efficient retrieval is possible. Assume we are indexing all words, and there is no stemming or preprocessing of words. 

\begin{tabular}{||c||c||c||c||}
    \hline
    Document 1 & Document 2 & Document 3 & Document 4    \\\hline
    the itsy bitsy   & down came       & out came      & dried up all            \\
    spider   &  the rain     & the sun      &     the rain          \\\hline\hline 
    Document 5 & Document 6 & Document 7 & Document 8    \\\hline
    the spider    & incy wincy      & went climbing     & washed poor             \\
    went up    &  spider     & up again      &     incy out          \\\hline\hline 
\end{tabular}

% answer:
%
% the -> 1, 2, 3, 4, 5
% itsy -> 1 
% bitsy -> 1
% spider -> 1, 5, 6
% down -> 2
% .... 

\vspace{9cm}
\item What documents would be retrieved given the query ``the AND spider"?
\end{enumerate}
% answer: doc 1 and 5

\newpage
\Question{\Large Information Extraction}{15}

Consider a machine that has the following text available (training data):

\begin{quotation}
\em Germany midfielder Florian Neuhaus scored on his national team debut but Turkey came back three times to earn a 3-3 draw in their friendly on Wednesday. Kenan Karaman slotted in a stoppage-time equaliser after Luca Waldschmidt fired in an 81st-minute volley that had put Germany 3-2 ahead. Turkey twice before had levelled, with both teams missing several regular players.

"We invited Turkey to score goals and again failed to hold on to victory," Germany captain Julian Draxler said.
Draxler's good finish from a Kai Havertz assist had put the hosts ahead on the stroke of halftime in front of 300 fans allowed in the stadium in Cologne, Germany.
\end{quotation}

and has to answer the following question:

\begin{quotation}
\em Which players scored a goal for Germany?
\end{quotation}

\begin{enumerate}
\item Draw a knowledge graph that the machine could construct from the data that would enable it to answer the question.

\vskip 7cm

\item Give two examples for co-reference in the text. Cite the referring expressions for each of the examples.

\end{enumerate}

\newpage
\section*{Extra Space}
\newpage
\end{document}

